---
title: "FORMATIVE ASSESSMENT 3"
author: "Lindsy Rossel C. Masicat"
date: "2024-02-25"
output: html_document
---

## Answer Exercises 7.1 items 2 and 7.
### Question Number 2
  A binary communication channel carries data as one of two sets of signals denoted by 0 and 1. Owing to noise, a transmitted 0 is sometimes received as a 1, and a transmitted 1 is sometimes received as a 0. For a given channel, it can be assumed that a transmitted 0 is correctly received with probability 0.95, and a transmitted 1 is correctly received with probability 0.75. Also, 70% of all messages are transmitted as a 0. If a signal is sent, determine the probability that:
  
  (a) a 1 was received;

  Based on the book, the solution will be

  Let R0 be the event that a zero is received. Let T0 be the event that a zero is transmitted.

  Let R1 be the event that a one is received. Let T1 be the event that a one is transmitted.
  
  the probability that a one was received, we note

  R1 = (T1 ∩ R1)∪(T0 ∩ R1)

  so the probability that one is received,

  P(R1) = P(T1)P(R1|T1) + P(T0)P(R1|T0)

  therefore,

  P(R1)=(0.3 × 0.75)+(0.7 × 0.05) = 0.225 + 0.035 = 0.26
  
  in R,

```{r R1}

  Received0_given_T0 <- 0.95
  Received1_given_T1 <- 0.75
  Received1_given_T0 <- 0.05
  Received0_given_T1 <- 0.25

  T0 <- 0.7
  T1 <- 0.3

  R1 <- ((T1*Received1_given_T1) + (T0*Received1_given_T0))

  cat("The probability that 1 is received is: ", R1, "\n")

```

  Thus, 26% is the probability that 1 is received.
  
  
(b) a 1 was transmitted given than a 1 was received. 

The solution will be,

  Let R0 be the event that a zero is received. Let T0 be the event that a zero is transmitted.
  Let R1 be the event that a one is received. Let T1 be the event that a one is transmitted.

  We use Bayes’ rule,

  P(T1|R1) = ((P(T1)P(R1|T1))/P(R1))

  Then, 

  P(T1)P(R1|T1) = 0.3 × 0.75 = 0.225

  Given that P(R1) = 0.26, based on the solution in letter A, we will now have

  P(T1|R1) = ((0.3 × 0.75)/0.26) = (0.225/0.26) = 0.865

  in R, 
  
```{r T1|RQ}
    Received0_given_T0 <- 0.95
    Received1_given_T1 <- 0.75
    Received1_given_T0 <- 0.05
    Received0_given_T1 <- 0.25

    T0 <- 0.7
    T1 <- 0.3

    R1 <- ((T1*Received1_given_T1) + (T0*Received1_given_T0))

    R1_given_T1 <- ((T1*Received1_given_T1)/R1)

    cat("The probability that 1 was transmitted given than a 1 was received: ",        R1_given_T1, "\n")
```
  
  Thus, 86.5% is the probability that 1 is transmitted given that 1 is received.


### Question Number 7
There are three employees working at an IT company: Jane, Amy, and Ava, doing 10%, 30%, and 60% of the programming, respectively. 8% of Jane’s work, 5% of Amy’s work, and just 1% of Ava‘s work is in error. What is the overall percentage of error? If a program is found with an error, who is the most likely person to have written it? 

To solve for the overall percentage error, we use total probability

  Assuming E to be the error of the work of the three employees,
  
  E = (Jane ∩ janeErrorWork)∪(Amy ∩ amyErrorWork)∪(Ava ∩ avaErrorWork)
  
  so,
  
  P(E) = P(Jane)P(E|Jane) + P(Amy)P(E|Amy) + P(Ava)P(E|Ava)
       = ((0.1*0.08) + (0.3*0.05) + (0.6*0.01))
       = 0.008 + 0.015 + 0.006
       = 0.029
       
  in R,
  
```{r overall error}
  janeProgram <- 0.1
  amyProgram <- 0.3
  avaProgram <- 0.6

  janeErrorWork <- 0.08
  amyErrorWork <- 0.05
  avaErrorWork <- 0.01
  
  totalError <- ((janeProgram*janeErrorWork) + (amyProgram*amyErrorWork) + (avaProgram*avaErrorWork))

  cat("The total probability of error: ", totalError, "\n")
```
  
  Thus, 2.9% is the total probability of overall error percentage.
  
  
  Now, to find the person who most likely have written the error, solve for posterior possibilities,
  
  P(J|E) = ((P(Jane)P(E|Jane))/P(E)) = ((0.1*0.08)/0.029) = 0.28
  P(A|E) = ((P(Amy)P(E|Amy))/P(E)) = ((0.3*0.05)/0.029) = 0.52
  P(AV|E) = ((P(Ava)P(E|Ava))/P(E)) = ((0.6*0.01)/0.029) = 0.21

  in R,

```{r most likely}

  janeError <- ((janeProgram*janeErrorWork)/totalError)
  cat("Jane's written error: ", janeError, "\n")

  amyError <- ((amyProgram*amyErrorWork)/totalError)
  cat("Amy's written error: ", amyError, "\n")
  
  avaError <- ((avaProgram*avaErrorWork)/totalError)
  cat("Ava's written error: ",  avaError, "\n")
 
  errorRate <- c(janeError, amyError, avaError)
  mostLikely <- which.max(errorRate)
  cat("The most likely person to have written the error is:", c("Jane", "Amy", "Ava")[mostLikely], "\n")
  
```

  Because P(A|E) is the largest, the person who most likely have written an error is Amy.